### 2022-10-14 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH presentation gold

\documentclass{beamer}
\usetheme{Boadilla}

\usepackage{graphicx}
\usepackage{subcaption}
% Change size of footnotes
\renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
\title{Beamer presentation}
%\subtitle{Using Beamer}
\author{Aitor Gonz\'alez}
\institute{TAGC/INSERM U1090}
\date{Internal Seminar: January 19, 2022}

% Add section slide
\AtBeginSection[]
{
\begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsection]
\end{frame}
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}

\titlepage

\end{frame}

\end{document}