### 2023-01-19 presentation_230120_internal_seminar/intern_sem_230120.tex
ENH

\documentclass{beamer}
\usetheme{Boadilla}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{csvsimple}

% Set relative path to all figures
\newcommand*{\floatRelativePath}{../out/gwas417/pval_5e-08/r2_0.1/kb_1000/window_1000000/75_50}%
% Footnote without marker: https://tex.stackexchange.com/questions/30720/footnote-without-a-marker
\newcommand\blfootnote[1]{%
    \begingroup
    \renewcommand\thefootnote{}\footnote{#1}%
    \addtocounter{footnote}{-1}%
    \endgroup
}
% Change size of footnotes
\renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
% reset footnote counter for each frame in beamer class
\AtBeginEnvironment{frame}{\setcounter{footnote}{0}}

\title{Analysis of pleiotropy of eQTLs}
\subtitle{Internal Seminar}
\author{Aitor Gonz\'alez}
\institute{Aix Marseille Univ, INSERM, TAGC}
\date{Jan. 20, 2023}

% Add section slide
\AtBeginSection[]
{
    \begin{frame}
        \frametitle{Table of Contents}
        \tableofcontents[currentsection]
    \end{frame}
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}

        \titlepage

    \end{frame}

    \section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    The associations to diseases are heterougeneously distributed in the genome, Watanabe & Posthuma
%    This is based on loci associated to multiple traits
    \begin{frame}
        \frametitle{Genome-wide association studies (GWAS)}

        \begin{itemize}
            \item Association test of effect allele on case individuals
            \item High-throughput genetics: 3000 unique loci for over 250 disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories
        \end{itemize}
%
        \vfill

%          However, how do we know if these variants are active in a tissue,
%         or whether these variants false positives due to linkage disequilibrium,
%         or how these variants act molecularly
        Questions:
%
        \begin{itemize}
            \item Correlation between variants. Which is the causal variant?
            \item Which cell types important for disease?
            \item In which tissues are these variants active?
        \end{itemize}
%
        \vfill
%
        Questions partially addressed by looking at eQTLs in different tissues

        \blfootnote{}
        \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Expression quantitative trait loci (eQTLs)}
%eQTLs are variants that modify gene expression in a given tissue.
%        eQTLs give a potential mechanism (the gene target) and a tissue.
        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}

            \end{column}
        \end{columns}
%
        \vfill
%
%        To my knowledge, it is unknown where are the most pleiotropice Eqtls in the genome.
%        This is important because this would give hints to putatitve pleiotropi genes (unknown or unknown) that are regulated by these pleiotropic eQTLs?
        Pleiotropy of eQTLs in genome is currently unknown

        \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Objectives}

        \begin{enumerate}
            \item To look for pleiotropic eQTLs in the genome
            \item To characterise their genomic regions, target genes and the cause of their pleiotropy
        \end{enumerate}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Strategy}

        \begin{enumerate}
            \item Annotate eQTLs with GWAS traits using colocalization analysis
            \item Annotate GWAS traits in categories
            \item Aggregate trait categories of eQTLs: Pleiotropic eQTLs
            \item Aggregate pleiotropic eQTLs in genome regions
            \item Analyse pleiotropic eQTLs and regions
        \end{enumerate}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Annotation of GWAS variants with gene and tissue targets using colocalization analysis}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4bc.png}
                \end{center}
            \end{column}
        \end{columns}

        \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Data and Algorithms}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{itemize}
                    \item IEA OpenGWAS database\footnote{https://gwas.mrcieu.ac.uk}: 418 GWAS datasets
                    \item EBI eQTLs database\footnote{https://www.ebi.ac.uk/eqtl}: Processed eQTLs from 127 studies in human.
                    \item CRAN - Package coloc\footnote{https://cran.r-project.org/web/packages/coloc/index.html}: Colocalisation analysis of GWAS/eQTL with one causal variant
                \end{itemize}
            \end{column}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/pgen.1004383.g001.png}
                \end{center}
            \end{column}
        \end{columns}

%        \let\thefootnote\relax\footnotetext{$^1$https://gwas.mrcieu.ac.uk}
%        \let\thefootnote\relax\footnotetext{$^2$https://www.ebi.ac.uk/eqtl}
%        \let\thefootnote\relax\footnotetext{$^3$https://cran.r-project.org/web/packages/coloc/index.html}
    \end{frame}

    \section{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Large scale GWAS/eQTL colocalization analysis}

%In this project, I have carried a large colocalization analysis between ~300 GWAS traits and 127 eQTL studies belonging to different tissues and mainly primary cell types.
%In addition, by colocalizing statistically eQTLs and GWAS, we can to some extent, break, linkage disequilibrium

        \begin{itemize}
            \item Colocalization between 293 GWAS and 127 eQTL studies
            \item 138e3 colocalized variants with probability $\geq$ 0.75
            \item Percentage of GWAS loci explained by eQTL: 0$\%$ - 77$\%$ (Average 33$\%$)
        \end{itemize}

        \begin{figure}[!]
            \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/coloc_web.png}
        \end{figure}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Web portail}

%These data can be useful to interpret disease variants
%        Therefore I have developed a small web portail

        TODO

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Do GWAS traits cluster coherently?}

%        To evaluate the coherence of the colocalization data, I have computed distances between traits based on the Spearman correlation of the betas of the eqtl-gene-tissue

        GWAS distance based on correlation of beta of colocalized eQTL gene and tissue

        \begin{figure}[!]
            \includegraphics[width=0.5\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
        \end{figure}

        \begin{itemize}
            \item Categories are coherent
            \item Help categorize GWAS traits
        \end{itemize}

%        In this heatmap, we can see that similar traits cluster together
%One difficulty for analysing pleiotropy is to distinguish similar traits for really different traits.
%In the previous heatmap, we have seen that traits cluster around coherent categories
%To look for real pleiotropic eQTLs I have manually annotated traits in different categories, such as cardiovascular disease, allergic, autoimmune, ....

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{frame}
%\frametitle{Diseases?}
%
%\begin{table}[!tbp]
%\centering
%\scriptsize
%\hline
%\csvreader[separator=tab,
%tabular=ccrrp{0.4\textwidth},
%head,
%table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
%]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
%{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
%\hline
%%
%\vspace{15pt}
%%
%\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
%\end{table}
%
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{How are pleiotropic eQTLs defined?}

        \begin{itemize}
            \item Uniformly annotate GWAS traits based on clustering and EBI ontologies
            \item Aggregate colocalised eQTLs based on the number of trait categories
            \item Analyse eQTLs in groups based on the number of trait categories
        \end{itemize}

    \end{frame}

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Winner (Most pleiotroic) eQTLs?}

        \begin{itemize}
            \item Example eQTLs involved in four GWAS categories in different cytobands
            \item The representative gene is the eQTL gene with the highest PubMed publication count
        \end{itemize}

% full size table is table
        \begin{table}[!tbp]
            \centering
            \footnotesize
            \csvreader[separator=tab,
            tabular=crclcp{0.4\textwidth},
            head,
            table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Known eQTL gene & \bfseries GWAS Categories\\\hline,
            ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
                {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
            \vspace{15pt}\label{tab:pleitropic_variants}
        \end{table}

    \end{frame}

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Winner (Most pleiotroic) regions?}

% full size table is table
        \begin{table}[!tbp]
            \centering
            \footnotesize
%\hline
            \csvreader[
                separator=tab,
                tabular=p{0.005\textwidth}p{0.05\textwidth}p{0.05\textwidth}p{0.05\textwidth}p{0.05\textwidth}p{0.5\textwidth},
                head,
                table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Known eQTL gene & \bfseries GWAS Categories\\\hline,
            ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
            {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
%\hline
%
            \vspace{15pt}
            \caption{Pleiotropic regions involving 6 or more GWAS classes. These regions were built with a sliding window of 1e5 nt. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
        \end{table}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{frame}
%    \frametitle{How are pleiotropic variants distributed?}
%
%    \begin{columns}
%        \begin{column}{0.32\textwidth}
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
%            \end{center}
%        \end{column}
%        \begin{column}{0.32\textwidth}
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
%            \end{center}
%        \end{column}
%        \begin{column}{0.32\textwidth}
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
%            \end{center}
%        \end{column}
%    \end{columns}
%
%    \begin{center}
%        \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/chr5_131912097_132802472.png}
%    \end{center}
%
%    \small
%    \begin{itemize}
%        \item The cytokine locus: chr5:131,912,097-132,802,472. Genes IRF1, IL3, IL4, IL5, ...
%        \item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
%        \item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
%    \end{itemize}
%    \normalsize
%
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Where are the pleiotropic variants?\footnote{https://www.ensembl.org/vep}}

        \begin{center}
            \includegraphics[width=0.7\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
        \end{center}

        In the transcripts

%        \let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Do pleiotropic variants associate to more eQTL genes?}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=0.7\textwidth]{../presentation_230120_gold2022_paris/fig/model_pleio_egenes.png}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Do pleiotropic variants associate to more eQTL tissues?}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=0.7\textwidth]{../presentation_230120_gold2022_paris/fig/model_pleio_etissues.png}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Do pleiotropic variants bind more transcription factors?\footnote{https://remap.univ-amu.fr}}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
                \end{center}
            \end{column}
        \end{columns}

%        \let\thefootnote\relax\footnotetext{$^1$https://remap.univ-amu.fr}

    \end{frame}

    \section{Conclusions, perspective and acknowledgements} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Conclusion}

        \begin{itemize}
            \item A list of pleiotrtA large set of causal GWAS variants annotated with target genes and tissues
            \item A list of pleiotropic variants and regions
        \end{itemize}

        Outcomes interesting for the GOLD population genotypes
%
        \begin{itemize}
            \item A large set of causal GWAS variants annotated with target genes and tissues
            \item A list of pleiotropic variants and regions
        \end{itemize}
%
        \vfill
%
        What are the gene regulatory properties underlying pleiotropy between complex traits?
%
        \begin{itemize}
            \item Variant pleiotropy is correlated in regions
            \item Enrichment of variants upstream (incl. 5'UTR) and downstream of genes (incl. 3'UTR)
            \item Enrichment of transcription factor and CRM binding
            \item Increased number of associated egenes and etissues
        \end{itemize}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Perspectives}

        \begin{itemize}
            \item Causality of eQTL genes on GWAS
            \item Study loci in more
        \end{itemize}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Acknowledgements}

        \begin{itemize}
            \item L\'eopoldine Lecerf (M1)
            \item P Paul, P Rihet, M Michel, S Marquet, S Spicuglia
        \end{itemize}
%
        \vfill
%
        Funding
%
        \begin{itemize}
            \item Institut Cancer et Immunologie - Aix-Marseille Univ.
            \item Agence nationale de la recherche (ANR)
            \item GOLD
        \end{itemize}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{frame}
%    \frametitle{Do pleiotropic variants associate to more traits (after fixing egene-etissue)?}
%
%    \begin{columns}
%        \begin{column}{0.5\textwidth}
%            \begin{center}
%                \includegraphics[width=0.7\textwidth]{../presentation_230120_gold2022_paris/fig/model_pleio_gwas.png}
%            \end{center}
%        \end{column}
%        \begin{column}{0.5\textwidth}  %%<--- here
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
%            \end{center}
%        \end{column}
%    \end{columns}
%
%\end{frame}

\end{document}