### 2023-02-18 ms/ms06_backmatter.tex
MS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

    \section*{Acknowledgements}%% if any
    Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
%
    I thank Sandrine Marquet, Marie Michel, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and Léopoldine Lecerf for a preliminary version of the colocalization pipeline developed during her internship.

%    \section*{Funding}%% if any
%    Text for this section\ldots

%    \section*{Abbreviations}%% if any
%    Text for this section\ldots

    \section*{Availability of data and materials}%% if any
    Text for this section\ldots

%    \section*{Ethics approval and consent to participate}%% if any
%    Text for this section\ldots

    \section*{Competing interests}
    The author declare that he has no competing interests.

%    \section*{Consent for publication}%% if any
%    Text for this section\ldots

    \section*{Authors' contributions}
    Aitor González: Conceptualization, Methodology, Software, Validation, Formal analysis, Resources, Data Curation, Visualization, Writing - original draft, Writing - review \& editing.

%    \section*{Authors' information}%% if any
%    Text for this section\ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
    \bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
    \bibliography{ms_pleiotropy}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

    \section*{Figures}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 1: Histogram of the percentage of the GWAS explained leading variants
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
        I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
            because this locus is not included in the colocalization pipeline.}
        \label{fig:hist_perc_tophits_eqtl_excl_mhc}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 2: Trait distance matrix
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.}
        \label{fig:trait_clusters}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
    \begin{figure}[h!]
        %
        \caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes
            (\textbf{b}) and eQTL tissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Spearman correlation between the counts of GWAS classes, eQTL genes and eQTL biosamples.}
        \label{fig:correlation_counts}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 5: VEP consequences
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Odds ratio of the variants annotated with given variant effect
        predictor (VEP) consequence and given GWAS class count compared with non
        pleiotropic variants with GWAS class count one. The significance is given for the
        q-value with False discovery rate 5\%.}
        \label{fig:vep_consequence}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 6: beta and pval
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{TODO beta and p-val}
        \label{fig:beta_pval}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Mean number of eQTL genes per eQTL-biosample pair (\textbf{a}) and mean number of biosamples per eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 8: TF count per GWAS class count
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Analysis of transcription factor binding.
        \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
        \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.}
        \label{fig:freq_tf_per_variant}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
    \section*{Tables}

    \input{ms07_tables}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \section*{Additional Files}
    \subsection*{Additional file 1: Table S1.}
    Classification of eQTL tissues and cell types. The first six were downloaded from the EBI eQTL repository.
    The 7th column "etissue\_category\_term" is used here to compute tissue diversity.

    \subsection*{Additional file 2: Table S2.}
    Metadata and classification of GWAS.

    \subsection*{Additional file 3: Table S3.}
    Percentage of tophits per GWAS that colocalized with at least one eQTL.

    \subsection*{Additional file 4: Table S4.}
    Count and list of GWAS phenotypes, egenes and etissues for each eQTL/GWAS variant.

    \subsection*{Additional file 5: Table S5.}
    Count and list of GWAS phenotypes, egene symbols and ENSEMBL IDs and etissue classes for each pleiotropic region.

\end{backmatter}