### 2023-02-26 ms/ms06_backmatter.tex
MS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

    \section*{Acknowledgements}%% if any
    Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
%
    I thank Sandrine Marquet, Marie Michel, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and Léopoldine Lecerf for a preliminary version of the colocalization pipeline developed during her internship.

%    \section*{Funding}%% if any
%    Text for this section\ldots

%    \section*{Abbreviations}%% if any
%    Text for this section\ldots

    \section*{Availability of data and materials}%% if any
    Text for this section\ldots

%    \section*{Ethics approval and consent to participate}%% if any
%    Text for this section\ldots

    \section*{Competing interests}
    The author declare that he has no competing interests.

%    \section*{Consent for publication}%% if any
%    Text for this section\ldots

    \section*{Authors' contributions}
    Aitor González: Conceptualization, Methodology, Software, Validation, Formal analysis, Resources, Data Curation, Visualization, Writing - original draft, Writing - review \& editing.

%    \section*{Authors' information}%% if any
%    Text for this section\ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
    \bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
    \bibliography{ms_pleiotropy}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

    \section*{Figures}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 1: Histogram of the percentage of the GWAS explained leading variants
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
        I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
            because this locus is not included in the colocalization pipeline.}
        \label{fig:1}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 2: Trait distance matrix
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.}
        \label{fig:2}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
% Fig 5: VEP consequences
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
    \begin{figure}[!ht]
        %
        \caption{Distribution of GWAS category counts, correlation and ENSEMBL variant effect prediction.
        \textbf{a-c} Distribution of GWAS category, eQTL gene and eQTL biosample counts.
        Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes (\textbf{b}) and eQTL tissues (\textbf{c}).
        \textbf{d} Spearman correlation between the counts of GWAS categories, eQTL genes and eQTL biosamples.
        \textbf{e} Odds ratio of the eQTLs annotated with the given ENSEMBL variant predictor
        for eQTLs association withe the given GWAS category count versus the eQTLs with a single GWAS category count.
        The significance is given for the q-value with False discovery rate 5\%.
        }
        \label{fig:3}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 6: beta and pval
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{eQTL and GWAS beta and p-values and allele frequencies for different GWAS category counts.
        \textbf{a,b} Average eQTL and GWAS beta values.
        \textbf{c,d} Average eQTL and GWAS p-values.
        \textbf{e} Average alternative allele frequency in the European population from the 1000 Genomes database.
        }
        \label{fig:4}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{Mean number of eQTL genes per eQTL-biosample pair (\textbf{a}) and mean number of biosamples per eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 8: TF count per GWAS class count
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{Analysis of transcription factor binding.
        \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
        \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.}
        \label{fig:freq_tf_per_variant}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
    \section*{Tables}

    \input{ms07_tables}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \section*{Additional Files}
    \subsection*{Additional file 1: Table S1.}
    Classification of eQTL tissues and cell types. The first six were downloaded from the EBI eQTL repository.
    The 7th column "etissue\_category\_term" is used here to compute tissue diversity.

    \subsection*{Additional file 2: Table S2.}
    Metadata and classification of GWAS.

    \subsection*{Additional file 3: Table S3.}
    Percentage of tophits per GWAS that colocalized with at least one eQTL.

    \subsection*{Additional file 4: Table S4.}
    Count and list of GWAS phenotypes, egenes and etissues for each eQTL/GWAS variant.

    \subsection*{Additional file 5: Table S5.}
    Count and list of GWAS phenotypes, egene symbols and ENSEMBL IDs and etissue classes for each pleiotropic region.

\end{backmatter}